Verifying Spatial Properties of Array Computations

### 2017-04-12 oopsla17/oopsla17.tex
add call to sloccount in VIEW mode

@@

\subsection{Semantic model}
\label{sec:semantics}

\noindent
We define a simple set-theoretic denotational model of the semantics
of our specification language. This has a number of purposes:
(1) it serves to explain the meaning of our specifications
(2) it is used in the inference and checking algorithms
(\Cref{sec:analysis}); (3) justifies an equational theory for
specifications in the next section; and (4) can be used to guide
correct implementations.

As an initial informal intuition, consider the following simple
stencil computation.
%
\begin{minted}{fortran}
do i = 1, n
  b(i) = a(i,0) + a(i+1,0)
end do
\end{minted}
%
This is modelled by a vector of integer sets: $\langle{\{0, 1\},
  \mathbb{Z}\}}\rangle$ describing that in the first dimension
the array is read from at offsets of $0$ and $1$ from an induction
variable, and in the second dimension the array reads are
unconstrained. \dnote{Don't like this explanation.}

we model neighbourhood array subscripts
in the context of a loop by vectors of sets

The model domain is over vectors of sets 
which we call \emph{index schemes}.

\begin{defn}[Extended integers]
  We define the set \zinf{} as
  $\mathbb{Z}$ extended with $\infty$ and $-\infty$. For any $a$ in
  \zinf{}, we have $-\infty \leq a \leq \infty$. The resulting set is
  a total order with top and bottom elements by $-\infty$ and
  $\infty$.
\end{defn}

\begin{defn}[Index scheme]
  An \emph{index scheme} is a vector of extended integer sets,
  that is, an index scheme of size $n$ is a value
  $\mathcal{P}(\zinf)^{n}$.
\end{defn}

>>>>

We used Wheeler's \emph{SLOCCount} to get a count of the physical
source lines (excluding comments and blank
lines)~\cite{wheeler2001sloccount}; including them our corpus
totals

@@

\section{Semantic model}
\label{sec:model}

\noindent
We define a simple set-theoretic model of array access patterns
which serves as both a denotational model of the semantics
of our specification language and an abstract interpretation
domain for source code. This model has a number of uses: 
(1) it serves to explain the meaning of our specifications
(2) it is used in the inference and checking algorithms
(\Cref{sec:analysis}); (3) justifies an equational theory for
specifications in the next section; and (4) can be used to guide
correct implementations.

The model is defined over vectors of sets of integers which we call
\emph{index schemes}.  As an initial informal example, consider the
following simple stencil computation:
%
\begin{minted}{fortran}
do i = 1, n
  b(i) = a(i,0) + a(i+1,0)
end do
\end{minted}
%
The access pattern on array \term{a}, relative
to induction variable \term{i}, is captured
by a vector of length 2 containing integer sets $\langle{\{0, 1\},
  \mathbb{Z}\}}\rangle$. This describes that, in the first dimension,
the array is read at offsets of $0$ and $1$ from an induction
variable. In the second dimension, the index is unconstrained
as it is a constant.

Index schemes form a lattice which provides a rich set of equations
and properties, which we exploit. We first set up the 
domain of the model (\S\ref{sec:domain}), using it to define a semantics
for our specification language (\S\ref{sec:semantics}) and then as
the target for an abstract interpretation on imperative
code (\S\ref{sec:fromcode}).

\subsection{Lattice model of regions}
\label{sec:domain}

\begin{defn}[Extended integers]
  We define the set \zinf{} as
  $\mathbb{Z}$ extended with $\infty$ and $-\infty$. For any $a$ in
  \zinf{}, we have $-\infty \leq a \leq \infty$. The resulting set is
  a total order with top and bottom elements $-\infty$ and
  $\infty$ respectively.
\end{defn}

\begin{defn}[Index scheme]
  An $N$-dimensional \emph{index scheme} is a vector of length $N$ of
  extended integer sets, \ie{}, a member of $\mathcal{P}(\zinf)^{N}$.
  An alternate (equivalent) view is of a $N$-times finite Cartesian
  product on subsets of $\zinf$.  We use $S, T, U$ to denote index
  schemes.

Index schemes can be \emph{projected} in the $i^{\textit{th}}$ dimension by
  $\pi_i : {\mathcal{P}(\zinf{})}^N \to \mathcal{P}(\zinf{})$ where $1 \leq i
  \leq N$. For an index scheme $I$, we refer to $\pi_i(S)$ as the
  \emph{$i^{th}$ component} of $S$.
\end{defn}

%\begin{theorem}
%$N$-dimensional index schemes form a lattice where join and m
%\end{theorem}

\begin{lem}{}\label{lem:vector-intersect}
  Intersection distributes over $N$-dimensional index schemes. 
  That is, for index schemes $I, J \in \mathcal{P}(\zinf)^N$ then:
%
  \begin{equation*}
    I \cap J = \prod_{i = 1}^{N} \pi_i(I) \cap \pi_i(J)
  \end{equation*}
\end{lem}

 Union does not distribute over $N$-dimensional index schemes--
 however, a more restricted property holds.

\begin{lem}{}\label{lem:vector-union}
Let $I, J$ be $N$-dimensional index schemes such that
$\pi_i(I) = \pi_i(J)$ for all $1 \leq i \leq N$ apart 
from some dimension $k$ where $\pi_k(I) \neq \pi_k(J)$,
then:
%
  \begin{equation*}
    I \; \cup \; J
    =
    \pi_1(I) \times \cdots \times
    (\pi_k(I) \cup \pi_k(J)) \times \cdots \times
    \pi_N(I)
  \end{equation*}
%
\end{lem}
%
\paragraph{Intervals}

We define an extended notion of closed interval on

\begin{defn}{(\emph{Interval subspace})}
  An interval subspace is a finite Cartesian product of intervals on \zinf{},
  denoted by the set $\textit{Interval}^N$ for a product of $N$ intervals. An
  interval subspace is an offset subspace.
\end{defn}

\subsection{Denotational semantics for specifications}
\label{sec:semantics}

\subsection{Abstract interpretation}
\label{sec:fromcode}
